---
title: "Estimating feeding kernel parameters"
output: html_notebook
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(tidyverse)
library(bbmle)
library(mizer)
```

We load our params object
```{r}
params <- readRDS("params_hake_kernel.rds")
```
We specify the latin.name we are interested in. Due to differences in spelling 
this is different from the list of species in the params object. Instead
it was supplied by Silvia:
```{r}
#select <- params@species_params$latin.name
select <- c("Aspitrigla cuculus",
"Lophius budegassa",
"Merluccius merluccius", 
"Arnoglossus imperialis",
"Micromesistius poutassou",
"Trachurus trachurus",
"Trisopterus minutus")
```

# Load data
Here we will use the Barnes dataset. If you don't have it, you can download it
with
```{r eval=FALSE, include=FALSE}
download.file("http://esapubs.org/archive/ecol/E089/051/Predator_and_prey_body_sizes_in_marine_food_webs_vsn4.txt", destfile = "barnes.csv")
```

We load the entire Barnes dataset and then select only the part we are
interested in. In particular we only keep the species we are interested in
and among those only those with more than 10 observations.
```{r}
stomach_all <-
    read_tsv("barnes.csv",
             na = c("", "n/a"),
             guess_max = 10000)

stomach <- stomach_all %>% 
    filter(Predator %in% select) %>% 
    # you could alternatively select by common name
    # filter(`Predator common name` %in% selected) %>% 
    transmute(latin.name = Predator,
              Nprey = 1,
              wpredator = `SI predator mass`,
              wprey = `SI prey mass`,
              l = log(wpredator / wprey)) %>% 
    # ignore prey that are larger than the predator
    filter(l > 0) %>% 
    group_by(latin.name) %>% 
    filter(n() > 10) %>% 
    mutate(weight_numbers = Nprey / sum(Nprey),
           weight_biomass = Nprey * wprey / sum(Nprey * wprey))
stomach
```
This leaves the following species:
```{r}
unique(stomach$latin.name)
```

  
# Bin the data
```{r}
no_bins <- 30  # Number of bins
binsize <- (max(stomach$l) - min(stomach$l)) / (no_bins - 1)
breaks <- seq(min(stomach$l) - binsize/2,
              by = binsize, length.out = no_bins + 1)
binned_stomach <- stomach %>% 
    # bin data
    mutate(cut = cut(l, breaks = breaks, right = FALSE,
                     labels = FALSE)) %>% 
    group_by(latin.name, cut) %>% 
    summarise(Numbers = sum(Nprey), 
              Biomass = sum(Nprey * wprey)) %>% 
    # normalise
    mutate(Numbers = Numbers / sum(Numbers) / binsize,
           Biomass = Biomass / sum(Biomass) / binsize)  %>%
    # column for predator/prey size ratio
    mutate(l = map_dbl(cut, function(idx) breaks[idx] + binsize/2)) %>% 
    gather(key = "Type", value = "Density", Numbers, Biomass)
```

# Normal distribution
Next we fit normal densities:
```{r}
grid <- seq(0, max(stomach$l), length = 100)
normaldens <- plyr::ddply(stomach, "latin.name", function(df) {
  data.frame( 
    l = grid,
    density = dnorm(grid, mean(df$l), sd(df$l))
  )
})

ggplot(stomach) +
    geom_density(aes(l, weight = weight_numbers), fill = "#00BFC4") +
    facet_wrap(~latin.name, scales = "free_y", ncol = 4) +
    xlab("Log of predator/prey mass ratio")  +
    geom_line(aes(l, density), data = normaldens,
                  colour = "blue")
```
Next we look at the biomass distribution. We plot the normal density arising
from the above fits on top of the observed biomass distribution.
```{r}
grid <- seq(-2, max(stomach$l), length = 100)
shifted_normaldens <- plyr::ddply(stomach, "latin.name", function(df) {
  data.frame( 
    l = grid,
    density = dnorm(grid, mean(df$l) - sd(df$l)^2, sd(df$l))
  )
})

ggplot(stomach) +
    geom_density(aes(l, weight = weight_biomass), fill = "#F8766D") +
    facet_wrap(~latin.name, scales = "free_y", ncol = 4) +
    xlab("Log of predator/prey mass ratio") +
    ylab("Biomass density") +
    geom_line(aes(l, density), data = shifted_normaldens,
                  colour = "red")
```
If we are happy with the fits, then we would set the following species parameters:
```{r}
lambda <- 2.05
sp <- stomach %>% 
    group_by(latin.name) %>% 
    summarize(sigma = sd(l),
              beta = mean(l) + (4/3 - lambda) * sigma^2)
sp
```



# Truncated exponential
We choose
$$
f_l(l) \propto \frac{\exp(\alpha\ l)}
{\left(1+e^{u_l(l_l - l)}\right)
\left(1+e^{u_r(l - l_r)}\right)}.
$$
```{r}
fl <- function(l, alpha, ll, ul, lr, ur) {
  dl <- ll - l
  dr <- l - lr
  fl <- exp(alpha * l) /
    (1 + exp(ul * dl)) /
    (1 + exp(ur * dr))
}
dtexp <- function(l, alpha, ll, ul, lr, ur) {
  d <- fl(l, alpha, ll, ul, lr, ur) /
    integrate(fl, 0, 30, alpha = alpha, 
              ll = ll, ul = ul, lr = lr, ur = ur)$value
  if (any(d <= 0)) {
    stop("The density contains non-positive values when",
         " alpha = ", alpha, " ll = ", ll, " ul = ", ul,
         " lr = ", lr, " ur = ", ur)
  }
  return(d)
}
```
We estimate the 5 parameters by maximum likelihood estimation, which in this
case can only be done numerically.
```{r}
mle_texp <- function(df) {
  loglik <- function(alpha, ll, ul, lr, ur) {
    L <- dtexp(df$l, alpha, ll, ul, lr, ur)
    - sum(log(L) * df$weight_numbers)
  }
  mle2(loglik, start = list(
    alpha = -0.5,
    ll = min(df$l),
    lr = max(df$l),
    ul = 5,
    ur = 5))
}
est <- stomach %>% 
  #filter(latin.name %in% select[c(1:3, 5:7)]) %>% 
  group_modify(~ broom::tidy(mle_texp(.x))) %>% 
  select(latin.name, term, estimate) %>% 
  spread(term, estimate)
est
```

```{r fig.height=15}
grid = seq(-2, max(stomach$l), length.out = 200)
texpdens <- plyr::ddply(est, "latin.name", function(df) {
  data.frame(
    l = grid,
    Numbers = dtexp(grid, df$alpha, df$ll, df$ul, df$lr, df$ur),
    Biomass = dtexp(grid, df$alpha - 1, df$ll, df$ul, df$lr, df$ur)
  )}) %>% 
  gather(Type, Density, Numbers, Biomass)

ggplot(binned_stomach) +
  geom_col(aes(l, Density, fill = Type)) +
  facet_grid(latin.name ~ Type, scales = "free_y") +
    xlab("Log of predator/prey mass ratio") +
    geom_line(aes(l, Density, colour = Type), data = texpdens)
```
```{r}
est
```

```{r}
lambda <- params@lambda
sp <- est %>% 
    transmute(kernel_exp = alpha + 4/3 - lambda,
              kernel_l_l = ll,
              kernel_u_l = ul,
              kernel_l_r = lr,
              kernel_u_r = ur)
sp
```
Let’s take a look at the resulting kernels
```{r}
grid = seq(-1, max(stomach$l) + 1, length.out = 200)
phi <- plyr::ddply(sp, "latin.name", function(df) {
  data.frame(
    l = grid,
    y = power_law_pred_kernel(exp(grid),
                              kernel_exp = df$kernel_exp,
                              kernel_l_l = df$kernel_l_l,
                              kernel_l_r = df$kernel_l_r,
                              kernel_u_l = df$kernel_u_l,
                              kernel_u_r = df$kernel_u_r)
  )})

ggplot(binned_stomach) +
  facet_wrap(~latin.name, scales = "free_y") +
  xlab("Log of predator/prey mass ratio") +
  geom_line(aes(l, y), size = 3, data = phi) +
  theme(strip.text = element_text(size = 12))
```

A dictionary to match the latin names to the common names in the params object.
```{r}
dict <- c("Merluccius merluccius" = "Hake", 
          "Aspitrigla cuculus" = "Gurnards", 
          "Lophius budegassa" = "Angler fish", 
          "Trisopterus minutus" = "Poor cod")
```

Put the parameters into the params object
```{r}
params@species_params$kernel_exp <- NA
params@species_params$kernel_l_l <- NA
params@species_params$kernel_u_l <- NA
params@species_params$kernel_l_r <- NA
params@species_params$kernel_u_r <- NA
for (i in seq_along(dict)) {
  common.name <- dict[i]
  idx <- which(params@species_params$species == common.name)
  params@species_params$pred_kernel_type[idx] = "power_law"
  params@species_params$kernel_exp[idx] <- sp$kernel_exp[i]
  params@species_params$kernel_l_l[idx] <- sp$kernel_l_l[i]
  params@species_params$kernel_l_r[idx] <- sp$kernel_l_r[i]
  params@species_params$kernel_u_l[idx] <- sp$kernel_u_l[i]
  params@species_params$kernel_u_r[idx] <- sp$kernel_u_r[i]
}
params <- setPredKernel(params)
```
```{r}
params <- tuneParams(params)
save(params, file = "params_barnes_kernels.rds")
```



