---
title: "Untitled"
output: html_document
---

We assume that the basic metabolic rate $R_b$ scales allometrically 
$$
R_b(M)=a\ M^{b}.
$$
We follow the paper by Clarke et.al. who measure $R_b$ in mmol oxygen per hour.
They give the exponent as
```{r}
b <- 0.8
```

Fishbase gives the metabolic rate for a Dourada of $M=78.3 g$ as 203mg/kg/h. To 
convert that to mmol oxygen per fish per hour we have to divide by the weight of
one mmol oxygen in mg (32) and multiply by the weight of the fish in kg
(M/1000)
```{r}
M <- 78.3
R_b <- 203 / 32 * (M / 1000)
R_b
```
Solving the metabolic scaling equation for $a$ gives
$$
a = R_b(M) / M^{b}
$$
```{r}
a <- R_b / M^b
a
```
This is the metabolic rate of a 1g Dourada, measured in mmol oxygen per hour.
We compare this to the average value for $a$ given by Clarke et.al
```{r}
exp(-5.43)
```
They differ by almost one order of magnitude!

We convert the mmol of oxygen into the weight of carbon needed to convert all
the oxygen into CO_2. Because one mmol of oxygen combines with one mmol of carbon
and one mmol of carbon weighs 12mg, the weight of carbon in g is
```{r}
carbonwt <- a * 12 / 1000
carbonwt
```
Let us assume that half of the dryweight is due to carbon:
```{r}
drywt <- carbonwt * 2
drywt
```
We convert from dry weight to wet weight using the conversion factor used by 
Erzini et.al.
```{r}
wetwt <- drywt/0.27
wetwt
```
Next we convert to a rate per year rather than per hour. There are 8760 hours per year
```{r}
wetwtyr <- wetwt * 8760
wetwtyr
```

```{r}
Ks <- wetwtyr/(1^0.7)
Ks
```


