---
title: "Untitled"
output: html_document
---

---
title: "Untitled"
output: html_document
---

We assume that the basic metabolic rate $R_b$ scales allometrically 
$$
R_b(M)=a\ M^{b}.
$$
We follow the paper by Clarke et.al. who measure $R_b$ in mmol oxygen per hour.
They give the exponent as
```{r}
b <- 0.8
```

Fishbase gives the metabolic rate for a Dourada of $M=78.3 g$ as 203mg/kg/h. To 
convert that to mmol oxygen per fish per hour we have to divide by the weight of
one mmol oxygen in mg (16) and multiply by the weight of the fish in kg
(M/1000)
```{r}
M_eel <- 99.80283
M_seabass <-192.875
M_sargo <- 22.875
M_wrasse <- 68
M_mullet <- 47
```
```{r}
o2_eel <- 112.437
o2_seabass <- 471.35
o2_sargo <- 396.6
o2_wrasse <-249.3
o2_mullet <-198.2
```
```{r}
R_b_eel <- o2_eel / 32 * (M_eel / 1000)
R_b_seabass <- o2_seabass / 32 * (M_seabass / 1000)
R_b_sargo <- o2_sargo / 32 * (M_sargo / 1000)
R_b_wrasse <- o2_wrasse / 32 * (M_wrasse / 1000)
R_b_mullet <- o2_mullet / 32 * (M_mullet / 1000)
```
```{r}
R_b_eel
R_b_seabass
R_b_sargo
R_b_wrasse
R_b_mullet
```
Solving the metabolic scaling equation for $a$ gives
$$
a = R_b(M) / M^{b}
$$
```{r}
a_eel <- R_b_eel / M_eel^b
a_seabass <- R_b_seabass / M_seabass^b
a_sargo <- R_b_sargo / M_sargo^b
a_wrasse <- R_b_wrasse / M_wrasse^b
a_mullet<- R_b_mullet / M_mullet^b


a_eel
a_seabass
a_sargo
a_wrasse
a_mullet
```
This is the metabolic rate of a 1g Dourada, measured in mmol oxygen per hour.
We compare this to the average value for $a$ given by Clarke et.al
```{r}
exp(-5.43)
```
They differ by almost one order of magnitude!

We convert the mmol of oxygen into the weight of carbon needed to convert all
the oxygen into CO_2. Because one mmol of oxygen combines with one mmol of carbon
and one mmol of carbon weighs 12mg, the weight of carbon in g is
```{r}
carbonwt_eel <- a_eel * 12 / 1000
carbonwt_seabass <- a_seabass * 12 / 1000
carbonwt_sargo <- a_sargo * 12 / 1000
carbonwt_wrasse <- a_wrasse * 12 / 1000
carbonwt_mullet <- a_mullet * 12 / 1000
carbonwt_eel
carbonwt_seabass
carbonwt_sargo
carbonwt_wrasse
carbonwt_mullet
```
Let us assume that half of the dryweight is due to carbon:
```{r}
drywt_eel <- carbonwt_eel * 2
drywt_seabass <- carbonwt_seabass * 2
drywt_sargo <- carbonwt_sargo * 2
drywt_wrasse <- carbonwt_wrasse * 2
drywt_mullet <- carbonwt_mullet * 2
drywt_eel
drywt_seabass
drywt_sargo
drywt_wrasse
drywt_mullet
```
We convert from dry weight to wet weight using the conversion factor used by 
Erzini et.al.
```{r}
wetwt_eel <- drywt_eel/0.27
wetwt_seabass <- drywt_seabass/0.27
wetwt_sargo <- drywt_sargo/0.27
wetwt_wrasse <- drywt_wrasse/0.27
wetwt_mullet <- drywt_mullet/0.27
wetwt_eel
wetwt_seabass
wetwt_sargo
wetwt_wrasse
wetwt_mullet
```
Next we convert to a rate per year rather than per hour. There are 8760 hours per year
```{r}
wetwt_eel_yr <- wetwt_eel*8760
wetwt_seabass_yr <- wetwt_seabass*8760
wetwt_sargo_yr <- wetwt_sargo*8760
wetwt_wrasse_yr <- wetwt_wrasse*8760
wetwt_mullet_yr <- wetwt_mullet*8760
wetwt_eel_yr
wetwt_seabass_yr
wetwt_sargo_yr
wetwt_wrasse_yr
wetwt_mullet_yr
```
```{r}
Ks_eel <- wetwt_eel_yr/(1^0.7)
Ks_seabass <- wetwt_seabass_yr/(1^0.7)
Ks_sargo <- wetwt_sargo_yr/(1^0.7)
Ks_wrasse <- wetwt_wrasse_yr/(1^0.7)
Ks_mullet <- wetwt_mullet_yr/(1^0.7)
Ks_eel
Ks_seabass
Ks_sargo
Ks_wrasse
Ks_mullet
```
```{r}

Ks_eel_Wmat <- wetwt_eel_yr/(220^0.7)
Ks_seabass_Wmat <- wetwt_seabass_yr/(716^0.7)
Ks_sargo_Wmat <- wetwt_sargo_yr/(334^0.7)
Ks_wrasse_Wmat <- wetwt_wrasse_yr/(37^0.7)
Ks_mullet_Wmat <- wetwt_mullet_yr/(586^0.7)
Ks_eel_Wmat
Ks_seabass_Wmat
Ks_sargo_Wmat
Ks_wrasse_Wmat
Ks_mullet_Wmat
```
```{r}
Ks_eel_Winf <- wetwt_eel_yr/(1643^0.7)
Ks_seabass_Winf <- wetwt_seabass_yr/(4623^0.7)
Ks_sargo_Winf <- wetwt_sargo_yr/(1950^0.7)
Ks_wrasse_Winf <- wetwt_wrasse_yr/(1330^0.7)
Ks_mullet_Winf <- wetwt_mullet_yr/(3368^0.7)
Ks_eel_Winf
Ks_seabass_Winf
Ks_sargo_Winf
Ks_wrasse_Winf
Ks_mullet_Winf

```



```{r}
Ks_dourada <- 11.81544
```
```{r}
species <-c("dourada","eel","seabass","sargo","wrasse","mullet")
Ks <-c(11.81544, 6.869726, 32.85491, 18.04789, 14.10671, 10.41657)
H <-c(48,12.8,73,42,24,30)
df <- data.frame(species,Ks,H)
df
```
```{r}
scatter.smooth(x=Ks, y=H, labels=row.names(species))
```


```{r}
cor(Ks, H)
```
```{r}
linearMod <- lm(Ks ~ H, data=df)  
print(linearMod)
```
```{r}
summary(linearMod)
```
Pr(>|t|) < 0.05    P-value < 0.05 
regression statistically significant 

```{r}
predict.sp <-c("choupa","baila","toadfish","pargo","garfish","sole", "mucharra", "Safia", "silverside", "redmullet", "salema", "goby", "anchovy", "sardie", "worms", "snails", "bivalve","crustacea")
Predict.H <-c(41,28,30,34,26,35,27,27,11,36,42,29,18,18.8,8.7,9,14,4.75)
predict.df <-data.frame(predict.sp,Predict.H)
predict.df
```
```{r}
predict.lm()
```
```{r}
predict.Ks <- predict(linearMod, labels=predict.sp)
predict.Ks
```
```{r}
species <-c("dourada","eel","seabass","sargo","wrasse","mullet")
Ks <-c(11.81544, 6.869726, 32.85491, 18.04789, 14.10671, 10.41657)
K_0 <-c(-0.55,-1.4,-0.63,-1,-0.6,-0.65)
df <- data.frame(species,Ks,K_0)
```
```{r}
scatter.smooth(x=Ks, y=K_0, labels=row.names(species))
```
```{r}
cor(Ks, K_vb)
```